```{r}
library(lattice)
library(xtable)
library(scales)

library(RSQLite)
library(RUnit)
library(gdata)
library(gridExtra)
library(RColorBrewer)
library(rentrez)
library(reshape2)
library(lubridate)
library(ggplot2)
library(gplots)
library(data.table)
setwd("~/temp/PdbMine/")

## GGplot default color scheme 
gg_color_hue <- function(n,luminance=65,chroma=100) {
  hues = seq(15, 375, length=n+1)
  hcl(h=hues, l=luminance, c=chroma)[1:n]
}

## A hack to make xtable work with dates 
xtable <- function(x, ...) {
   for (i in which(sapply(x, function(y) !all(is.na(match(c("POSIXt","Date"),class(y))))))) x[[i]] <- as.character(x[[i]])
   xtable::xtable(x, ...)
}

.simpleCap <- function(x) {
    s <- strsplit(x, " ")[[1]]
    paste(toupper(substring(s, 1,1)), tolower(substring(s, 2)),
          sep="", collapse=" ")
}

trim <- function (x) gsub("^\\s+|\\s+$", "", x)


```

We build the data table 
```{r}
m <- dbDriver("SQLite")
con <- dbConnect(m, dbname = "local-medline-pdb.db")
sql<-"SELECT pdbid,depodate, pmid,journal,
cast(earliestpublicdate as chard) as earliestpublicdate,
cast(receiveddate as char) as received,
cast(aheadofprintdate as char) as aheadofprint,
cast(accepteddate as char) as accepted,
cast(publisheddate as char) as published,
cast(earliestpublicationdate as char) as earliestpublication
FROM pdb INNER JOIN medline ON jrnlref=pmid 
WHERE 
    earliestpublicdate IS NOT NULL 
AND earliestpublicdate >=Datetime('1995-01-01')
AND (published IS NULL OR published < Datetime('2014-01-01'))
AND (accepted IS NULL OR accepted < Datetime('2014-01-01'))"
#AND (accepted IS NULL OR received IS NULL OR received<=accepted
merged<-dbGetQuery(con,sql)
merged<-unique(merged)
dim(merged)
length(unique(merged$pdbid))
length(unique(merged$pmid))
#save(merged,file="data/pdb.medline.RData")
```

We can then load it 
```{r}
load("data/pdb.medline.RData")
```

We recast the dates into R date type and build a data.table
```{r}
merged$earliestpublicdate <- as.Date(merged$earliestpublicdate)
merged$earliestpublication <- as.Date(merged$earliestpublication)
merged$depodate <- as.Date(merged$depodate)
merged$published <- as.Date(merged$published)
merged$received <- as.Date(merged$received)
merged$aheadofprint <- as.Date(merged$aheadofprint)
merged$accepted <- as.Date(merged$accepted)

# Using lubridate helpers
merged$pubyear <- as.factor(year(merged$earliestpublication))
merged$depoyear <- as.factor(year(merged$depodate))


merged$journal <- as.factor(merged$journal)
merged$pmid <- as.factor(merged$pmid)
pdb.medline <- data.table(merged)
pdb.medline[,has.history:=!is.na(received) | !is.na(accepted),by=pdbid]
#save(merged,pdb.medline,file="data/pdb.medline.RData")
```

We filter out outliers, coming from NSMB that are accepted before review 
```{r}
outliers<-pdb.medline[!is.na(accepted) & !is.na(received) & received>accepted,pmid]
pdb.medline <- pdb.medline[!(pmid %in% outliers)]

outliers <- pdb.medline[!is.na(accepted) & earliestpublication<accepted,pmid]
pdb.medline <- pdb.medline[!(pmid %in% outliers)]
save(merged,pdb.medline,file="data/pdb.medline.RData")
```


We rename some journals whose name changed over time 
```{r}
pdb.medline[journal=="Nat Struct Biol",journal:="Nat Struct Mol Biol"]
pdb.medline[journal=="Fold Des",journal:="Structure"] # Only 4 entries????

pdb.medline[journal=="Acta Crystallogr D Biol Crystallogr",journal:="Acta Crystallogr D"]
pdb.medline[journal=="Acta Crystallogr Sect F Struct Biol Cryst Commun",journal:="Acta Crystallogr F"]
save(merged,pdb.medline,file="data/pdb.medline.RData")
```
Categorize each pub into sloppy or late 
```{r}
pdb.medline[,time:="On time"]
pdb.medline[ as.integer(depodate - earliestpublication)>=2,time:="Late"]
pdb.medline[as.integer(depodate - earliestpublicdate)>=2 & time=="On time",time:="Sloppy"]
save(merged,pdb.medline,file="data/pdb.medline.RData")

# Old sloppy 
#pdb.medline[depodate> earliestpublicdate,.N,by=journal][order(N)]
```


* Do we have sloppy deposit for articles without publication history? (Accepted and/or received)
We shouldn't
```{r}
pdb.medline[time=="Sloppy" & (is.na(accepted) & is.na(received))]
```

* Do we have deposits that happened after publication that are neither sloppy or late? We shouldn't
```{r}
pdb.medline[time=="On time" & (depodate - earliestpublicdate>1)]
```
# Summary table 
```{r}
summary.per.journal<-pdb.medline[,list(N=.N,"With.History"=sum((!is.na(received) | !is.na(accepted)) ),"sloppy"=sum(time=="Sloppy"),"late"=sum(time=="Late")),by=journal]
summary.per.journal[,sloppy.ratio.min:=ifelse(With.History>0,sloppy/N,NA)]
summary.per.journal[,sloppy.ratio:=ifelse(With.History>0, sloppy/With.History,NA)]
summary.per.journal[,late.ratio:=late/N]
summary.per.journal<-head(summary.per.journal[order(N,decreasing=T)],n=50)
````

# Adding impact factor

# Generate the table with merged IF 
```{r}
load("data/impact.factor.2012.RData")

manual.journal.correction=new.env()
manual.journal.correction[["Haematologica"]]="Acta Haematol-basel"
manual.journal.correction[["J Comput Aided Mol Des"]]="J Comput Aid Mol Des"
manual.journal.correction[["Acta Crystallogr D Biol Crystallogr"]]="Acta Crystallogr D"
manual.journal.correction[["Acta Crystallogr Sect F Struct Biol Cryst Commun"]]="Acta Crystallogr F"
manual.journal.correction[["Biochemistry"]]="Biochemistry-us"
manual.journal.correction[["Epigenetics Chromatin"]]="Epigenet Chromatin"
manual.journal.correction[["Circ Cardiovasc Genet"]]="Circ-cardiovasc Gene"
manual.journal.correction[["N Engl J Med"]]="New Engl J Med"
manual.journal.correction[["Comp Biochem Physiol B Biochem Mol Biol"]]="Comp Biochem Phys B"
manual.journal.correction[["Angew Chem Int Ed Engl"]]="Angew Chem Int Edit"
manual.journal.correction[["Mol Biochem Parasitol"]]="Mol Biochem Parasit"
manual.journal.correction[["Antimicrob Agents Chemother"]]="Antimicrob Agents Ch"
manual.journal.correction[["Genes Dev"]]="Gene Dev"
manual.journal.correction[["Proc Natl Acad Sci U S A"]]="P Natl Acad Sci Usa"
manual.journal.correction[["Nat Struct Biol"]]="Nat Struct Mol Biol"
manual.journal.correction[["Eur J Biochem"]]="J Biochem"
manual.journal.correction[["Protein Pept Lett"]]="Protein Peptide Lett"
manual.journal.correction[["Chemistry"]]="Chem-eur J"
manual.journal.correction[["Biochim Biophys Acta"]]="Bba-proteins Proteom"
manual.journal.correction[["Appl Microbiol Biotechnol"]]="Appl Microbiol Biot"
manual.journal.correction[["Appl Environ Microbiol"]]="Appl Environ Microb"
manual.journal.correction[["Dalton Trans"]]="Dalton T"
manual.journal.correction[["Biochem Biophys Res Commun"]]="Biochem Bioph Res Co"
manual.journal.correction[["Chem Commun (Camb)"]]="Chem Commun"


impact.factor<-data.frame(impact.factor)
rownames(impact.factor)<-impact.factor$journal


match.journal.name<-function(j,approx=F,top1=F){
  
  if(j %in% ls(manual.journal.correction)){
    cat(j,"\n")
    return(manual.journal.correction[[j]])
  }
	# Exact match?
	if(j %in% rownames(impact.factor)){
		return(j)
	}
	# Exact case insensitive match?
	if(.simpleCap(j) %in% rownames(impact.factor)){
		return(.simpleCap(j))
	}
		
	j<-.simpleCap(j)
	if(approx){
		res<-rownames(impact.factor[agrep(j,impact.factor$journal,fixed=T,ignore.case=F),])
	}else{
		res<-rownames(impact.factor[grep(j,impact.factor$journal,fixed=T,ignore.case=F),])
	}
  
	if(top1){
		sub=impact.factor[res,]
		sorted<-rownames(sub[order(sub$Impact.Factor,decreasing=T),])
		return(sorted[1])
	}else{

		if(length(res)>=1){
			return(res)
		}else{
			return(character(0))
		}
 	}
}

match.journal.name("Nat Prot")

sapply(as.character(tail(pdb.medline[,.N,by=journal][order(N)][,journal],n=20)),match.journal.name)
```

We generate and display count data with annotated IF 
```{r}
library(xlsx)
summary.per.journal[,IF:=impact.factor[match.journal.name(as.character(journal)),"IF"],by=journal]
summary.per.journal
write.csv(summary.per.journal,quote=F,row.names=F,file="ratio.per.journal.csv")
write.xlsx(summary.per.journal,row.names=F,file="ratio.per.journal.xlsx")
```

# Plots
We plot the difference between depo and public as a function of depo, for the 20 journals with most pub
```{r}
pdb.medline.pop <- pdb.medline[depodate >= "2000-01-01" & depodate <= "2012-12-31"]
top.journals <-tail(pdb.medline.pop[,.N,by=journal][order(N),journal],n=20)

ggplot(pdb.medline.pop[journal %in% top.journals],aes(x=depodate,y=as.integer(earliestpublicdate-depodate),colour=journal))+geom_point()+geom_smooth()
ggplot(pdb.medline.pop[journal %in% top.journals],aes(x=depodate,y=as.integer(earliestpublicdate-depodate),colour=time))+geom_point()+geom_smooth()


ggplot(pdb.medline.pop[journal %in% top.journals],aes(x=depodate,y=published,colour=is.sloppy))+geom_point()+geom_smooth()
ggplot(pdb.medline.pop[journal %in% top.journals],aes(x=depodate,y=published,colour=is.late))+geom_point()+geom_smooth()

```


```{r}
pdb.medline.pop <- pdb.medline[depodate >= "1995-01-01" & depodate <= "2012-12-31"]
totPubs<-max(pdb.medline.pop[,.N,by=depoyear][,N])
g<-ggplot()+geom_bar(data=pdb.medline.pop,aes(x=depoyear,fill=time),position="fill")
g+geom_line(data=pdb.medline.pop[,.N,by=depoyear],aes(x=depoyear,y=N/totPubs,group=1),size=1,colour="grey20")+geom_point(data=pdb.medline.pop[,.N,by=depoyear],aes(x=depoyear,y=N/totPubs),size=2)
# ggplot(pdb.medline.pop[,.N,by=pubyear],aes(x=pubyear,y=N,group=1))+geom_line()
```


# Main figure 
If we filter by publication year instead of depoyear 
```{r}

# colors<-brewer_pal(type="div",palette="Set1")(9)[c(3,5,1)]
# 
# colors<-gg_color_hue(3,chroma=80,lum=65)[c(2,3,1)]
#  colors<-gg_color_hue(3)
# colors<-gg_color_hue(12,luminance=70)[c(5,2,1)] # Green orange red 
# colors<-gg_color_hue(20,luminance=70)[c(10,5,1)] 
colors<-gg_color_hue(3,chroma=90,luminance=60)[c(2,3,1)] # Tassos selection
# colors<-gg_color_hue(12,chroma=90,luminance=50)[c(5,2,1)] # Green orange red 

only.with.history <- F

if (only.with.history){
  pdb.medline.pop <- pdb.medline[ earliestpublicdate >= "1995-01-01" & earliestpublication <= "2012-12-31" & (!is.na(received) | !is.na(accepted))]
  # Merge entries before 2000
  pdb.medline.pop[pubyear=="1999",pubyear:="2000"]
  pdb.medline.pop[pubyear=="1998",pubyear:="2000"]
}else{
  pdb.medline.pop <- pdb.medline[ earliestpublicdate >= "1995-01-01" & earliestpublication <= "2012-12-31"]

}

pdb.medline.pop[,.N,by=pubyear][order(pubyear)]
pdb.medline.pop$time <- factor(pdb.medline.pop$time,levels=c("On time","Sloppy","Late"))

#write.xlsx(pdb.medline.pop,file="pdb.merged.with.medline.post.1995.all.entries.xlsx")

write.xlsx(pdb.medline.pop[,.N,by=list(pubyear,time)][order(pubyear)],file="late.sloppy.per.year.xlsx")

# Make a matrix like 
m1 <- pdb.medline.pop[,.N,by=list(pubyear,time)][order(pubyear)]
setnames(m1,"time","variable")
m2 <- pdb.medline.pop[,.N,by=list(pubyear,has.history)][has.history==T,list(pubyear,N)]
m2$variable="has.history"
m3 <- pdb.medline.pop[,.N,by=list(pubyear)]
m3$variable="total"

write.xlsx(acast(rbind(m1,m2,m3,use.names=T),variable~pubyear),file="categories.per.year.xlsx")

totPubs<-max(pdb.medline.pop[,.N,by=pubyear][,N])*1.05
g<-ggplot()+geom_bar(data=pdb.medline.pop,aes(x=pubyear,fill=time),position="fill")
g<-g+geom_line(data=pdb.medline.pop[,.N,by=pubyear],aes(x=pubyear,y=N/totPubs,group=1),size=1,colour="grey40")+geom_point(data=pdb.medline.pop[,.N,by=pubyear],aes(x=pubyear,y=N/totPubs),size=2,colour="grey20")
g<-g+scale_y_continuous(labels = percent_format())+xlab("Year of publication")+ylab("Proportion")
g<-g+geom_text(data=pdb.medline.pop[,.N,by=pubyear],aes(x=pubyear,y=N/totPubs,label=N),size=4.2,vjust=-1)
g<-g+scale_fill_manual(values=colors)

#+ggtitle(paste(colors,collapse="-"))

if(only.with.history){
  ggsave("proportion.late.sloppy.per.year.pdb.only.history.pdf",w=12,h=8,plot=g)
}else{
  ggsave("proportion.late.sloppy.per.year.pdb.pdf",w=12,h=8,plot=g)
}
print(g)

```

Facetted plot for counts and rates 
```{r}
# Only keep datum for which the review date is available ?

pdb.medline.pop <- pdb.medline[earliestpublicdate >= "1995-01-01" & earliestpublicdate <= "2012-12-31"]
pdb.medline.pop$panel<-"Ratio"
pdb.medline.pop$time <- factor(pdb.medline.pop$time,levels=c("On time","Sloppy","Late"))
counts <- pdb.medline.pop[,.N,by=pubyear]
counts$panel <- "Counts"
counts$time <- "All"
g<-ggplot()+facet_grid(panel~.,scale="free")
g<-g+geom_bar(data=pdb.medline.pop,aes(x=pubyear,fill=time),position="fill")
g<-g+geom_line(data=counts,aes(x=pubyear,y=N,group=1),size=1,colour="grey20")+geom_point(data=counts,aes(x=pubyear,y=N),size=2)
g
```
How many publications have gone public in 2012 but not published in 2012? This would explain the differences between selection based on earliestpublicdate and on published date 
```{r}
pdb.medline[earliestpublicdate <= "2012-12-31" & published>"2013-01-01"]
```

It's the 1.3k that we were missing? 
```{r}
pdb.medline[earliestpublicdate >= "1995-01-01" & earliestpublicdate <= "2012-12-31"][,.N,by=pubyear][order(pubyear)]

pdb.medline[published >= "1995-01-01" & published <= "2012-12-31"][,.N,by=pubyear][order(pubyear)]

```





Any association between IF and sloppy ?
```{r}
ggplot(summary.per.journal[N>10,],aes(x=IF,y=sloppy.ratio,size=N))+geom_point()+geom_smooth()
ggplot(summary.per.journal[N>10,],aes(x=IF,y=late.ratio,size=N))+geom_point()+geom_smooth()


```

* Do people publishing in Science are more in advance than people publishing in Nature? 
```{r}
ggplot(pdb.medline[journal %in% c("Nature","Science","Cell","Proteins","Proc Natl Acad Sci U S A") & earliestpublication>="2000-01-01"],aes(x=earliestpublication,y=as.integer(depodate-earliestpublication),group=journal, colour=journal,shape=time) )+geom_point()+geom_smooth(method="lm")+coord_cartesian(ylim=c(-300,100))


ggplot(pdb.medline[journal %in% summary.per.journal[N>1000,journal] & pubyear %in% c("2010","2011","2012")],aes(x=reorder(journal,summary.per.journal[order(sloppy.ratio)][journal]),y=as.integer(depodate-earliestpublication), fill=journal) )+geom_boxplot(notch=T)+coord_cartesian(ylim=c(-300,100))


```
Do the distribution of days in advance match the sloppy percent? 
```{r}
pdb.medline.with.IF=merge(pdb.medline,summary.per.journal,by="journal")

pdb.medline.with.IF$journal.by.sloppy <- factor(pdb.medline.with.IF$journal,levels=summary.per.journal[order(sloppy.ratio)][,journal])

ggplot(pdb.medline.with.IF[N>2000 & pubyear %in% c("2010","2011","2012")],aes(x=journal.by.sloppy,y=as.integer(depodate-earliestpublication), fill=IF) )+geom_boxplot(notch=T)+coord_cartesian(ylim=c(-300,100))
ggplot(pdb.medline.with.IF[N>1146 & pubyear %in% c("2010","2011","2012")],aes(x=journal.by.sloppy,y=as.integer(depodate-earliestpublication), fill=sloppy.ratio) )+geom_boxplot(notch=T)+coord_cartesian(ylim=c(-300,100))


ggplot(pdb.medline.with.IF,aes(x=IF,y=as.integer(depodate-earliestpublicdate),color=time))+geom_point()+geom_smooth(method="lm")+coord_cartesian(ylim=c(-250,250))+scale_x_continuous(lim=c(0,40))

ggplot(pdb.medline.with.IF,aes(x=earliestpublicdate,y=as.integer(depodate-earliestpublicdate),color=time))+geom_point()+geom_smooth(method="lm")+coord_cartesian(ylim=c(-250,250))

```

# Heatmap of sloppiness by journal and year
```{r}

summary.per.journal<- pdb.medline[!is.na(received),.N,by=journal][N>=20]
summary.per.journal[,IF:=impact.factor[match.journal.name(as.character(journal)),"IF"],by=journal]


top.journals <-tail(summary.per.journal[order(N),journal],n=70)


summary.per.journal.year<-pdb.medline[journal %in% top.journals & earliestpublication>="2002-01-01",list(N=.N,"With.History"=sum((!is.na(received) | !is.na(accepted)) ),"sloppy"=sum(time=="Sloppy"),"late"=sum(time=="Late")),by=list(journal,pubyear)]

summary.per.journal.year[,sloppy.ratio.min:=ifelse(With.History>0,sloppy/N,NA)]
summary.per.journal.year[,sloppy.ratio:=ifelse(With.History>0, sloppy/With.History,NA)]
summary.per.journal.year[,late.ratio:=late/N]
summary.per.journal.year[,IF:=impact.factor[match.journal.name(as.character(journal)),"IF"],by=journal]
#sloppy.per.journal.year <- acast(na.omit(summary.per.journal.year),journal~pubyear,value.var="late")

# Build the ordering
journals.by.IF <- summary.per.journal.year[,mean(IF),by="journal"][order(V1,na.last=F)][,journal]

summary.per.journal.year$journal <- factor(summary.per.journal.year$journal,levels=journals.by.IF)

write.xlsx(summary.per.journal.year,file="ratio.per.journal.and.year.xlsx",row.names=F)

colours=c("gold","white","red")
colours.breaks=c(0,0.5,1)

ggplot(na.omit(summary.per.journal.year[pubyear!="2013"]),aes(x=pubyear,y=journal,fill=sloppy.ratio,alpha=With.History>=10))+geom_tile()+geom_text(aes(label=paste(sloppy,With.History,sep="/")))+scale_fill_gradientn(colours=colours,values=colours.breaks)+scale_size_continuous(range=c(2,4))
# ggsave("heatmap.sloppy.per.year.per.journal.sorted.by.IF.pdf",width=20,height=12)
# 
# colours.breaks.late=c(0,0.05,0.4)
# ggplot(na.omit(summary.per.journal.year[pubyear!="2013"]),aes(x=pubyear,y=journal,alpha=N>=10,fill=late.ratio))+geom_tile()+geom_text(aes(label=paste(late,N,sep="/")))+scale_fill_gradientn(colours=colours,values=colours.breaks.late)+scale_size_continuous(range=c(2,4))

#summary.per.journal<-head(summary.per.journal[order(N,decreasing=T)],n=50)

```



## Heatmap with sloppy and late combined 
```{r}

colours = c("#00AA37","white","#E26D65")
colours=c("gold","white","red")
colours.breaks=c(0,0.5,1)


g<-ggplot(summary.per.journal.year[pubyear!="2013"],aes(x=pubyear,y=journal,fill=(sloppy+late)/N))+geom_tile()+geom_text(aes(label=paste(sloppy+late,N,sep="/")))+scale_fill_gradientn(colours=colours,values=colours.breaks)+scale_size_continuous(range=c(2,4))+xlab("Year of publication")+ylab("Journal")

ggsave("sloppy.and.late.pdb.per.year.and.journal.sorted.by.IF.heatmap.pdf",w=20,h=16,plot=g)
print(g)
```


## Summary per journal corresponding to the heatmap data 
```{r}
summary.per.journal<-pdb.medline[journal %in% top.journals,list(N=.N,"With.History"=sum((!is.na(received) | !is.na(accepted)) ),"sloppy"=sum(time=="Sloppy"),"late"=sum(time=="Late")),by=journal]
summary.per.journal[,sloppy.ratio.min:=ifelse(With.History>0,sloppy/N,NA)]
summary.per.journal[,sloppy.ratio:=ifelse(With.History>0, sloppy/With.History,NA)]
summary.per.journal[,late.ratio:=late/N]
summary.per.journal[,IF:=impact.factor[match.journal.name(as.character(journal)),"IF"],by=journal]
write.xlsx(summary.per.journal,file="ratio.per.journal.xlsx",row.names=F)
```

# Analysis of GEO Deposit time
```{r}
m <- dbDriver("SQLite")
con <- dbConnect(m, dbname = "local-medline-pdb.db")
sql<-"SELECT gse,submission_date, pubmed_id,pmid,journal,
cast(earliestpublicdate as chard) as earliestpublicdate,
cast(receiveddate as char) as received,
cast(aheadofprintdate as char) as aheadofprint,
cast(accepteddate as char) as accepted,
cast(publisheddate as char) as published,
cast(earliestpublicationdate as char) as earliestpublication
FROM gse INNER JOIN medline ON pubmed_id=pmid 
WHERE 
    earliestpublicdate IS NOT NULL 
AND earliestpublicdate >=Datetime('1995-01-01')"
#AND (published IS NULL OR published < Datetime('2013-01-01'))
#AND (accepted IS NULL OR accepted < Datetime('2013-01-01'))"
#AND (accepted IS NULL OR received IS NULL OR received<=accepted
merged.gse<-dbGetQuery(con,sql)
merged.gse<-unique(merged.gse)
dim(merged.gse)
length(unique(merged.gse$gse))
length(unique(merged.gse$pmid))
```

Cast the dates as appropriate
```{r}
merged.gse$earliestpublicdate <- as.Date(merged.gse$earliestpublicdate)
merged.gse$earliestpublication <- as.Date(merged.gse$earliestpublication)
merged.gse$submission_date <- as.Date(merged.gse$submission_date)
merged.gse$published <- as.Date(merged.gse$published)
merged.gse$received <- as.Date(merged.gse$received)
merged.gse$aheadofprint <- as.Date(merged.gse$aheadofprint)
merged.gse$accepted <- as.Date(merged.gse$accepted)

# Using lubridate helpers
merged.gse$pubyear <- as.factor(year(merged.gse$earliestpublication))
merged.gse$depoyear <- as.factor(year(merged.gse$submission_date))

gse.medline <- data.table(merged.gse)
gse.medline[,has.history:=!is.na(received) | !is.na(accepted),by=gse]
```
Classification
```{r}
gse.medline[,time:="On time"]
gse.medline[ as.integer(submission_date - earliestpublication)>=2,time:="Late"]
gse.medline[as.integer(submission_date - earliestpublicdate)>=2 & time=="On time",time:="Sloppy"]

# Checks 
gse.medline[time=="On time" & (submission_date - earliestpublicdate>1)]
gse.medline[time=="Sloppy" & (is.na(accepted) & is.na(received))]
```
Distribution per year
```{r}

colors<-gg_color_hue(3,chroma=90,luminance=60)[c(2,3,1)] # Tassos selection
# colors<-gg_color_hue(12,chroma=90,luminance=50)[c(5,2,1)] # Green orange red 
only.with.history <- T

if (only.with.history){
  gse.medline.pop <- gse.medline[ earliestpublicdate >= "1995-01-01" & earliestpublication <= "2012-12-31" & (!is.na(received) | !is.na(accepted))]
}else{
  gse.medline.pop <- gse.medline[ earliestpublicdate >= "1995-01-01" & earliestpublication <= "2012-12-31"]

}

gse.medline.pop[,.N,by=pubyear][order(pubyear)]
gse.medline.pop$time <- factor(gse.medline.pop$time,levels=c("On time","Sloppy","Late"))
totPubs<-max(gse.medline.pop[,.N,by=pubyear][,N])*1.05

g<-ggplot()+geom_bar(data=gse.medline.pop,aes(x=pubyear,fill=time),position="fill")

g<-g+geom_line(data=gse.medline.pop[,.N,by=pubyear],aes(x=pubyear,y=N/totPubs,group=1),size=1,colour="grey40")+geom_point(data=gse.medline.pop[,.N,by=pubyear],aes(x=pubyear,y=N/totPubs),size=2,colour="grey20")
g<-g+scale_y_continuous(labels = percent_format())+xlab("Year of publication")+ylab("Proportion")
g<-g+geom_text(data=gse.medline.pop[,.N,by=pubyear],aes(x=pubyear,y=N/totPubs,label=N),size=4.2,vjust=-1)
g+scale_fill_manual(values=colors)
#+ggtitle(paste(colors,collapse="-"))


if(only.with.history){
  ggsave("GEO.proportion.late.sloppy.per.year.pdb.only.history.pdf",w=12,h=8)
}else{
  ggsave("GEO.proportion.late.sloppy.per.year.pdb.pdf",w=12,h=8)
}



```

# Summary table 
```{r}
gse.summary.per.journal<-gse.medline[,list(N=.N,"With.History"=sum((!is.na(received) | !is.na(accepted)) ),"sloppy"=sum(time=="Sloppy"),"late"=sum(time=="Late")),by=journal]
gse.summary.per.journal[,sloppy.ratio.min:=ifelse(With.History>0,sloppy/N,NA)]
gse.summary.per.journal[,sloppy.ratio:=ifelse(With.History>0, sloppy/With.History,NA)]
gse.summary.per.journal[,late.ratio:=late/N]

gse.summary.per.journal[,IF:=impact.factor[match.journal.name(as.character(journal)),"IF"],by=journal]
summary.per.journal


gse.summary.per.journal<-head(gse.summary.per.journal[order(N,decreasing=T)],n=50)

write.xlsx(gse.summary.per.journal,row.names=F,file="GEO.ratio.per.journal.xlsx")

````



# Heatmap
```{r}
top.journals <-tail(gse.summary.per.journal[order(N),journal],n=70)

# Build the ordering
journals.by.IF <- gse.summary.per.journal[order(IF)][!is.na(IF)][,journal]

summary.per.journal.year<-gse.medline[journal %in% top.journals & earliestpublication>="2002-01-01",list(N=.N,"With.History"=sum((!is.na(received) | !is.na(accepted)) ),"sloppy"=sum(time=="Sloppy"),"late"=sum(time=="Late")),by=list(journal,pubyear)]
summary.per.journal.year[,sloppy.ratio.min:=ifelse(With.History>0,sloppy/N,NA)]
summary.per.journal.year[,sloppy.ratio:=ifelse(With.History>0, sloppy/With.History,NA)]
summary.per.journal.year[,late.ratio:=late/N]
summary.per.journal.year[,IF:=impact.factor[match.journal.name(as.character(journal)),"IF"],by=journal]
summary.per.journal.year$journal <- factor(summary.per.journal.year$journal,levels=journals.by.IF)
colours=c("gold","white","red")
colours.breaks=c(0,0.5,1)

ggplot(na.omit(summary.per.journal.year[pubyear!="2013"]),aes(x=pubyear,y=journal,fill=sloppy.ratio,alpha=With.History>=10))+geom_tile()+geom_text(aes(label=paste(sloppy,With.History,sep="/")))+scale_fill_gradientn(colours=colours,values=colours.breaks)+scale_size_continuous(range=c(2,4))

ggsave("GSE.heatmap.sloppy.per.year.per.journal.sorted.by.IF.pdf",width=20,height=12)


# Late 
colours.breaks.late=c(0,0.05,0.6)
ggplot(na.omit(summary.per.journal.year[pubyear!="2013"]),aes(x=pubyear,y=journal,alpha=N>=10,fill=late.ratio))+geom_tile()+geom_text(aes(label=paste(late,N,sep="/")))+scale_fill_gradientn(colours=colours,values=colours.breaks.late)+scale_size_continuous(range=c(2,4))

```





# During review, when does the structure gets deposited? 
I would expect an uniform distrib, since the authors are not aware of when a paper will be accepted 
```{r}

colors<-gg_color_hue(3,chroma=90,luminance=60)[c(2,3,1)] # Tassos selection

pdb.medline.pop <- pdb.medline[depodate >= "2000-01-01" & depodate <= "2012-12-31"]

pdb.medline.full.history <- pdb.medline.pop[has.history==T & !is.na(accepted) & !is.na(received) & accepted <= "2014-01-01"]


ggplot(pdb.medline[has.history==T & !is.na(accepted) &  time =="Sloppy"],aes(x=as.integer(depodate-accepted)))+geom_bar()

ggplot(pdb.medline[has.history==T & !is.na(accepted) &  time =="Sloppy"],aes(x=as.integer(depodate-accepted)))+geom_bar(position="stack",binwidth=5)+coord_cartesian(xlim=c(-100,+50))


ggplot(pdb.medline.pop[has.history==T & !is.na(accepted) &  time =="Sloppy"],aes(x=as.integer(depodate-accepted)))+geom_bar(position="stack",binwidth=1)+coord_cartesian(xlim=c(-100,+50))


ggplot(pdb.medline.pop[has.history==T & !is.na(accepted) &  time =="Sloppy" & journal %in% top.journals],aes(fill=journal,x=as.integer(depodate-accepted)))+geom_bar(position="stack",binwidth=5)+coord_cartesian(xlim=c(-100,+50))


ggplot(pdb.medline.pop[has.history==T & !is.na(accepted) &  time =="Sloppy" & journal %in% top.journals],aes(fill=journal,x=as.integer(depodate-accepted)))+geom_bar(position="stack",binwidth=2)+coord_cartesian(xlim=c(-100,+50))


ggplot(pdb.medline.pop[has.history==T & !is.na(accepted) &  time =="Sloppy" & journal %in% top.journals],aes(x=as.integer(depodate-accepted)))+geom_bar(position="stack",binwidth=2)+coord_cartesian(xlim=c(-100,+50))

ggplot(pdb.medline.pop[has.history==T & !is.na(accepted),list(time,delay=as.integer(depodate-accepted))],aes(fill=time,x=delay))+geom_bar(position="stack",binwidth=30)+coord_cartesian(xlim=c(-100,+50))


ggplot(pdb.medline.pop[has.history==T & !is.na(accepted),list(time,delay=as.integer(depodate-accepted))],aes(fill=time,x=delay))+geom_bar(position="stack",binwidth=2)+coord_cartesian(xlim=c(-100,+50))


# We accumulate anything before 20 days 
pdb.medline.full.history.sloppy.trimmed <- pdb.medline.full.history[!is.na(received)]
pdb.medline.full.history.sloppy.trimmed[,delay:=as.integer(depodate-accepted)]
pdb.medline.full.history.sloppy.trimmed[,pub.delay:=as.integer(depodate-earliestpublication)]

pdb.medline.full.history.sloppy.trimmed<-pdb.medline.full.history.sloppy.trimmed[!is.na(delay) ]


THR_DAY=20L
pdb.medline.full.history.sloppy.trimmed[delay<=-THR_DAY,delay:=-THR_DAY]
pdb.medline.full.history.sloppy.trimmed[delay>=THR_DAY,delay:=THR_DAY]

pdb.medline.full.history.sloppy.trimmed$time <- factor(pdb.medline.full.history.sloppy.trimmed$time,levels=c("On time","Sloppy","Late"))

ggplot(pdb.medline.full.history.sloppy.trimmed[-THR_DAY<=delay & delay<=THR_DAY],aes(fill=time,x=delay))+geom_bar(position="stack",binwidth=2)+geom_vline(x=2,linetype="dashed",color="firebrick")+scale_fill_manual(values=colors)+scale_x_continuous(limits=c(-THR_DAY,THR_DAY))

ggplot(pdb.medline.full.history.sloppy.trimmed[-THR_DAY<=delay & delay<=THR_DAY  & time=="Sloppy"],aes(fill=time,x=delay))+geom_bar(position="stack",binwidth=2)+geom_vline(x=2,linetype="dashed",color="firebrick")+scale_fill_manual(values=colors)


ggplot(pdb.medline.full.history.sloppy.trimmed[-THR_DAY<=pub.delay & pub.delay<=THR_DAY],aes(fill=time,x=pub.delay))+geom_bar(position="stack",binwidth=2)+geom_vline(y=0)+scale_fill_manual(values=colors)

```



Generate final pub ready version figures: Distribution of delays between deposition and paper acceptance 
```{r}
THR_DAY=50
colors<-gg_color_hue(3,chroma=90,luminance=60)[c(2,3,1)] # Tassos selection
pdb.medline.pop <- pdb.medline[depodate >= "2000-01-01" & depodate <= "2012-12-31"]
pdb.medline.full.history <- pdb.medline.pop[has.history==T & !is.na(accepted) & !is.na(received) & accepted <= "2014-01-01"]
pdb.medline.full.history[,delay:=as.integer(depodate-accepted)]
pdb.medline.full.history[,pub.delay:=as.integer(depodate-earliestpublication)]
pdb.medline.full.history$time=factor(pdb.medline.full.history$time,levels=c("On time","Sloppy","Late"))

g<-ggplot(pdb.medline.full.history[-THR_DAY<=pub.delay & pub.delay<=THR_DAY],aes(fill=time,x=pub.delay))+geom_bar(position="stack",binwidth=2)+geom_vline(x=2,linetype="dashed",color="firebrick")+scale_fill_manual(values=colors)

g<-ggplot(pdb.medline.full.history[-THR_DAY<=delay & delay<=THR_DAY],aes(fill=time,x=delay))+geom_bar(position="stack",binwidth=2)+geom_vline(x=2,linetype="dashed",color="firebrick")+scale_fill_manual(values=colors)


g<-g+xlab(paste("Number of days between day of structure deposition and day of paper acceptance (",nrow(pdb.medline.full.history)," data points)"))
ggsave(file="distribution.delays.deposit.acceptance.pdf",w=12,h=8,plot=g)
print(g)
```


Generate binned data in xls file
GGplot bin the data by excluding excluding the right value and including the left value 

```{r}
bins <- c(min(pdb.medline.full.history$delay),seq(-50,50,2),max(pdb.medline.full.history$delay),1)
bins <- c(min(pdb.medline.full.history$delay),seq(-50,50,2),max(pdb.medline.full.history$delay))
pdb.medline.full.history$bin<-cut(pdb.medline.full.history$delay,bins,right=F,include=T)
pdb.medline.full.history.binned<-pdb.medline.full.history[,.N,by=list(bin,time)][order(bin)]
pdb.medline.full.history.binned$time=factor(pdb.medline.full.history.binned$time,levels=c("On time","Sloppy","Late"))
ggplot(pdb.medline.full.history.binned[bin!="[-3.7e+03,-50)"],aes(x=bin,y=N,fill=time))+geom_bar(stat="identity")+scale_fill_manual(values=colors)
write.xlsx(pdb.medline.full.history.binned,file="distribution.delays.deposit.acceptance.xlsx",row.names=F)
write.xlsx(acast(pdb.medline.full.history.binned,bin~time,value.var="N"),file="distribution.delays.deposit.acceptance.xlsx",sheet="casted",append=T,showNA=F)
```



Problem is that this statistics depends on the delta between received and accepted dates. If we plot relative time, one for entry, we get 
** NOT FINISHED YET ** 

```{r}
pdb.medline.full.history.sloppy[,total.review.time:= as.integer(accepted-received)]
pdb.medline.full.history.sloppy[,relative.depo:= as.integer(depodate-accepted)/total.review.time]
```


# Four classes figure

## Data preparation 
```{r}
pdb.medline.4 <-  pdb.medline[ earliestpublicdate >= "1995-01-01" & earliestpublication <= "2012-12-31"]
pdb.medline.4[,time:="On time?"]
pdb.medline.4[ as.integer(depodate - earliestpublication)>=2,time:="Late"]
pdb.medline.4[as.integer(depodate - earliestpublicdate)>=2 & time=="On time?",time:="Sloppy"]
pdb.medline.4[as.integer(depodate - earliestpublicdate)<=2 & has.history==T &  time=="On time?",time:="Sure on time"]

pdb.medline.4$time <- factor(pdb.medline.4$time,levels=c("Late","On time?","Sloppy","Sure on time"),ordered=T)

pdb.medline.4.by.year=pdb.medline.4[,.N,list(time,pubyear)]
# Add, for each year, late and maybe 
pdb.medline.4.by.year.under<-data.table(dcast(pdb.medline.4.by.year,pubyear~time,value.var="N"))
setnames(pdb.medline.4.by.year.under,colnames(pdb.medline.4.by.year.under),make.names(colnames(pdb.medline.4.by.year.under)))

pdb.medline.4.by.year.under[,acc.sloppy:=On.time.+Late]
pdb.medline.4.by.year.under[,acc.sure:=On.time.+Late+Sloppy]

pdb.medline.4.by.year.under[,tot:=Late+On.time.+Sloppy+Sure.on.time,by=pubyear]
melt(pdb.medline.4.by.year.under)

pdb.medline.4.by.year.under[,with.history:=Sloppy+Sure.on.time,by=pubyear]


pdb.medline.4.by.year.under[,slop.label:= paste("atop(",Sloppy,",\"",sprintf("%d%%",as.integer(floor(Sloppy/with.history*100))),"\")",sep="")]

pdb.medline.4.by.year.under[,on.time.label:= paste("atop(",Sure.on.time,",\"",sprintf("%d%%",as.integer(ceiling(Sure.on.time/with.history*100))),"\")",sep="")]

pdb.medline.4.by.year.under[,late.label:=Late]

```

Facetted plot
```{r}
colors<-gg_color_hue(3,chroma=90,luminance=60)[c(2,3,1)] # Tassos selection
g1<-ggplot(pdb.medline.4,aes(x=pubyear,fill=time))+geom_bar(position="fill")
g2<-  ggplot(pdb.medline.4,aes(x=pubyear,fill=time))+geom_bar(position="stack")+scale_fill_manual(values=c(colors[3],"grey60",colors))
grid.arrange(g1,g2)

```
Combined plot
```{r}

geom_size=15
text_size=3.5
colors.4 <- colors
g<-ggplot()+geom_bar(data=pdb.medline.4,aes(x=pubyear,fill=time),position="stack")+scale_fill_manual(values=c(colors[3],rgb(200,200,200,alpha=200,maxColorValue=255),colors[2],colors[1]))
#Chnage data frame
#g<-g %+% na.omit(pdb.medline.4.by.year.under)
g<-g %+% pdb.medline.4.by.year.under
#g<-g+geom_line(aes(x=pubyear,y=acc.sloppy*(1+Sloppy/tot),label=Sloppy,group=1))
# Add Sure on time labeling 
g<-g+geom_point(aes(x=pubyear,y=acc.sure*(1+Sure.on.time/tot)),shape=22,size=geom_size,fill="#33DD7A")
g<-g+geom_text(aes(x=pubyear,y=acc.sure*(1+Sure.on.time/tot),label=on.time.label),size=text_size,parse=T)

#add sloppy labels
g<-g+geom_point(aes(x=pubyear,y=acc.sloppy*(1+Sloppy/tot)),shape=22,size=geom_size,fill="#6CB1FF")
g<-g+geom_text(aes(x=pubyear,y=acc.sloppy*(1+Sloppy/tot), label=slop.label),size=text_size,parse=T)

#add late labels
g<-g+geom_point(aes(x=pubyear,y=100),shape=22,size=geom_size*0.7,fill="#F37E76")
g<-g+geom_text(aes(x=pubyear,y=100, label=late.label),size=text_size,parse=T)
g<-g+xlab("Year of publication")
#ggsave(file="proportion.late.sloppy.per.year.pdb.four.classes.pdf",w=12,h=8,plot=g)
print(g)
```



